---
title: "data_exploration"
author: "Weiheng Zhang"
date: "2021/11/28"
output: github_document
---

```{r message = FALSE}
library(tidyverse)
library(lubridate)
library(dplyr)
library(p8105.datasets)
library(leaflet)
library(MASS)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Data cleaning

```{r message = FALSE, warning = FALSE}
cdi =
  read_csv("./data/cdi.csv") %>% 
  mutate(
    cty = as.factor(cty),
    state = as.factor(state),
    region = factor(region, levels = c("1", "2", "3", "4"),
                    labels = c("Northeast", "North_Central", "South", "West")),
    CRM_1000 = 1000*crimes/pop #每1000人中的犯罪数量(county level)
  ) %>% 
  relocate(CRM_1000)

cdi
```

## Descriptive Statistics.
```{r}
summary(cdi)
```

```{r}
sapply(cdi, function(x) sum(is.na(x)))
```
No missing values were found.



## Identify counties with unusual crime rates.
```{r}
upper = quantile(cdi$CRM_1000, 0.75)
lower = quantile(cdi$CRM_1000, 0.25)
IQR = upper - lower

cdi %>% 
  filter(CRM_1000 > upper + 1.5*IQR,
         CRM_1000 > lower - 1.5*IQR) %>% 
  dplyr::select(cty, CRM_1000) %>%
  knitr::kable(digits = 2)


cdi %>% 
  ggplot(aes(x = "", y = CRM_1000))+
  geom_boxplot() + 
  labs(title = "Counties' CRM_1000 Values",)
```


## Group by states and check for outliers again.
```{r}
cdi_state = 
  cdi %>% 
  group_by(state) %>% 
  summarise(state_pop = sum(pop),
            state_crimes = sum(crimes)) %>% 
  mutate(state_CRM_1000 = 1000*state_crimes/state_pop) %>% 
  relocate(state_CRM_1000)

cdi_state
```

```{r}
upper = quantile(cdi_state$state_CRM_1000, 0.75)
lower = quantile(cdi_state$state_CRM_1000, 0.25)
IQR = upper - lower

cdi_state %>% 
  filter(state_CRM_1000 > upper + 1.5*IQR,
         state_CRM_1000 > lower - 1.5*IQR) %>% 
  dplyr::select(state, state_CRM_1000) %>%
  knitr::kable(digits = 2)


cdi_state %>% 
  ggplot(aes(x = "", y = state_CRM_1000))+
  geom_boxplot() + 
  labs(title = "States' CRM_1000 Values",)
```
  
Surprisingly, if we look at the CRM_1000 at state level, no outlier was found.







## Check for transformation

```{r}
# fit multivariate model
mult.fit1 = lm(CRM_1000 ~ poverty + region + state + area + pop + pop18 + pop65 + hsgrad + bagrad + unemp + pcincome + totalinc, data = cdi) 

# check diagnostics
plot(mult.fit1)
boxcox(mult.fit1)
```

## a = 1/2. Perform sqrt transformation.

```{r}
cdi = 
  cdi %>% 
  mutate(sqrt_CRM_1000 = sqrt(CRM_1000))

mult.fit2 = lm(sqrt_CRM_1000 ~ poverty + region + state + area + pop + pop18 + pop65 + hsgrad + bagrad + unemp + pcincome + totalinc, data = cdi) 

# check diagnostics
summary(mult.fit2)
plot(mult.fit2)
boxcox(mult.fit2) 
```






## Initial Exploration of correlation between counties' CRM_100 and all variables, With untransformed outcome (CRM_1000)
## 好像除了poverty, region和state之外都看不出什么明显关联（捂脸）

```{r}
cdi %>% ggplot(aes(x = area, y = CRM_1000)) + geom_point()
cdi %>% ggplot(aes(x = pop, y = CRM_1000)) + geom_point()
cdi %>% ggplot(aes(x = pop18, y = CRM_1000)) + geom_point()
cdi %>% ggplot(aes(x = pop65, y = CRM_1000)) + geom_point()
cdi %>% ggplot(aes(x = docs, y = CRM_1000)) + geom_point()
cdi %>% ggplot(aes(x = beds, y = CRM_1000)) + geom_point()
cdi %>% ggplot(aes(x = hsgrad, y = CRM_1000)) + geom_point()
cdi %>% ggplot(aes(x = bagrad, y = CRM_1000)) + geom_point()
cdi %>% ggplot(aes(x = poverty, y = CRM_1000)) + geom_point()
cdi %>% ggplot(aes(x = unemp, y = CRM_1000)) + geom_point()
cdi %>% ggplot(aes(x = pcincome, y = CRM_1000)) + geom_point()
cdi %>% ggplot(aes(x = totalinc, y = CRM_1000)) + geom_point()
cdi %>% ggplot(aes(x = region, y = CRM_1000)) + geom_boxplot()
cdi %>% ggplot(aes(y = state, x = CRM_1000)) + geom_boxplot()
```










